A Survey on Ordered Binary Decision Diagrams

### 2015-09-15 paper.tex
Revise Operations intro and Restrict

As the name suggests, \textit{Reduced} Ordered Binary Decision Diagrams are a 
minimized representation of BDDs. Furthermore, the decision variables are ordered
consistently. More formally, for any node $n$ and their respective children $m_1$
and $m_2$, a total ordering is imposed, i.e. $var(n) < var(m_1)$ and
$var(n) < var(m_2)$, respectively, where $var(x)$ denotes the label of node
$x$~\cite{BRYANT86, BRYANT92}.

This \textit{variable ordering} affects the efficiency of the symbolic
manipulation because different variable orderings, in general, produce different
OBDDs. In the worst case, this will result in exponential complexity for bad
orderings. Nevertheless, as will be described in Section~\ref{sec:limitations},
there exist classes of functions for which the OBDD complexity is exponential
regardless of which variable ordering is chosen.

Before discussing the reduction procedure, some definitions are necessary. Here
the original definitions of Bryant~\cite{BRYANT86, BRYANT91} are reused.

A Boolean function can be represented as an OBDD, i.e. a rooted, directed,
acyclic graph (DAG) with a set of nodes $V$ as well as a set of edges $E$.

@@

For each constant function, there exists only a single terminal node representing
it, e.g. for Boolean functions, only two terminal nodes exist, representing
\textbf{0} and \textbf{1}, respectively. Hence, when transforming a given BDD,
eliminate all but one terminal node for each constant function and redirect all
edges accordingly, i.e. to the one remaining node for each constant function.

There may be duplicate nonterminal nodes present in a given BDD. Formally, a node
$d$ is a duplicate if $var(d) = var(n) \land lo(d) = lo(n) \land hi(d) = hi(n)$
holds for any other node $n$. If this is the case, $d$ can be eliminated and all
edges are redirected to $n$.

The third and last transformation rule deals with completely redundant nonterminal
nodes. Intuitively, this rule should be straightforward: any node having the same
child on all outgoing edges is redundant because the actual value of the variable
does not change the result of the function. More formally, a nonterminal node $n$
is redundant if and only if $lo(n) = hi(n)$. Then, all incoming edges of $n$ can
be redirected to $lo(n)$ and $n$ can be deleted afterwards.

@@

OBDDs have one big advantage over other representations: their representation is
\textit{canonical}. As a consequence, important tests can be performed on an OBDD
in constant time, i.e. satisfiability and tautology tests. Moreover, OBDDs enable
other tests, such as variable independence and functional equivalence tests, to
be performed efficiently. \tbr

@@

Bryant's~\cite{BRYANT86, BRYANT92} main focus was to define algorithms performing
symbolic operations on graphs representing Boolean functions. The most basic
operation is \textit{Apply} which applys a given operation $op$, e.g. \texttt{AND},
to two given functions $f_1$ and $f_2$, and returns the result of $f_1\;op\;f_2$.
Another operation is \textit{Restrict}, restricting a given variable $x$ to a
given (constant) value $c$, denoted $f|_{x\leftarrow c}$. Furthermore, the
operations \textit{Compose} and \textit{Satisfy} can be defined as graph
algorithms.

@@

A variable $x_i$, argument of a function $f$, can be restricted to a given value
$c$, denoted by $f|_{x_i\leftarrow c}$, as follows.

Starting from the root $n$ of the OBDD representation of $f$ and traversing the
OBDD depth-first, every node is checked for $var(n) = x_i$. This is done recursively
on the given graph, generating the resultant graph This node represents
the variable to be restricted. Now, there are two cases for $c$ to distinguish:
\begin{enumerate}
    \item{$c=0$: The incoming edge of $n$ is redirected to $lo(n)$.}
    \item{$c=1$: The incoming edge of $n$ is redirected to $hi(n)$.}
\end{enumerate}

@@

This procedure has linear time complexity in the number of nodes, i.e. $O(|V|)$,
assuming an efficient implementation (see
Section~\ref{sec:implementational-aspects} for details)~\cite{BRYANT92}. 
\tbr

>>>>

As the name suggests, \textit{Reduced} Ordered Binary Decision Diagrams are a 
minimized representation of BDDs. Furthermore, the decision variables are ordered
consistently. More formally, for any node $n$ and their respective children $m_1$
and $m_2$, a total ordering is imposed, i.e. $var(n) < var(m_1)$ and
$var(n) < var(m_2)$, respectively, where $var(x)$ denotes the label (variable)
of node $x$~\cite{BRYANT86, BRYANT92}.

This \textit{variable ordering} affects the efficiency of the symbolic
manipulation because different variable orderings, in general, produce different
OBDDs. In the worst case, this will result in exponential complexity for bad
orderings. Nevertheless, as will be described in Section~\ref{sec:limitations},
there exist classes of functions for which the OBDD complexity is exponential
regardless of the variable ordering chosen.

Before discussing the reduction procedure, some definitions are necessary. Here,
the original definitions of Bryant~\cite{BRYANT86, BRYANT91} are reused.

A Boolean function can be represented as an OBDD, i.e. a rooted, directed,
acyclic graph (DAG), with a set of nodes $V$ and a set of edges $E$.

@@

For each constant function, there exists only a single terminal node representing
it, i.e. for Boolean functions, only two terminal nodes exist, representing
\textbf{0} and \textbf{1}, respectively. Hence, when transforming a given BDD,
eliminate all but one terminal node for each constant function and redirect all
edges accordingly, i.e. to the one remaining node for each constant function.

There may be duplicate nonterminal nodes present in a given BDD. Formally, a node
$d$ is a duplicate if $var(d) = var(n) \land lo(d) = lo(n) \land hi(d) = hi(n)$
holds for any other node $n$. If this is the case, $d$ can be eliminated and all
edges are redirected to $n$ accordingly.

The third and last transformation rule deals with completely redundant nonterminal
nodes. Intuitively, this rule should be straightforward: any node having the same
child on both outgoing edges is redundant because the actual value of the variable
does not change the result of the function. More formally, a nonterminal node $n$
is redundant if and only if $lo(n) = hi(n)$. Then, all incoming edges of $n$ can
be redirected to $lo(n)$ and $n$ can be deleted afterwards.

@@

OBDDs have one big advantage over other representations: their representation is
\textit{canonical}. As a consequence, important tests can be performed on an OBDD
in constant time, i.e. satisfiability and tautology tests. Moreover, OBDDs enable
other tests, such as variable independence and functional equivalence tests, to
be performed efficiently.

@@

Bryant's~\cite{BRYANT86, BRYANT92} main focus was to define algorithms performing
symbolic operations on graphs representing Boolean functions. The most basic
operation is \textit{Apply} which applys a given operation $op$, e.g. \texttt{AND},
to two given functions $f_1$ and $f_2$, and returns the result of $f_1\;op\;f_2$.
Another operation is \textit{Restrict}, restricting a given variable $x$ to a
given (constant) value $c$, denoted $f|_{x=c}$. Furthermore, the
operations \textit{Compose} and \textit{Satisfy} can be defined as graph
algorithms.

@@

A variable $x_i$, argument of a function $f$, can be restricted to a given value
$c$, denoted by $f|_{x_i=c}$, as follows.

Starting from the root $n$ of the OBDD representation of $f$ and traversing the
OBDD depth-first, every node is checked for $var(n) = x_i$. This is done recursively
on the given graph, generating the resultant graph. This node represents
the variable to be restricted. Now, there are two cases for $c$ to be
distinguished.

\clearpage
\begin{enumerate}
    \item{$c=0$: The incoming edge of $n$ is redirected to $lo(n)$.}
    \item{$c=1$: The incoming edge of $n$ is redirected to $hi(n)$.}
\end{enumerate}

@@

This procedure has linear time complexity in the number of nodes, i.e. $O(|V|)$,
assuming an efficient implementation (see
Section~\ref{sec:implementational-aspects} for details)~\cite{BRYANT92}.
